---
myimage1: "../img/section/Choropleth150.png"
myimage2: "../img/section/MapHexbin150.png"
myimage3: "../img/section/Cartogram150.png"
mytitle: "Visualizing a set of geographic coordinates"
output:
  html_document:
      template: template_story.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: TRUE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---

<br><br>
<div class="mycontent">




<div class = "row">
<div class = "col-md-8">
<br><br><br>This document gives a few suggestions to analyse a dataset composed by a list of GPS coordinates<br>It considers the geographic position of about 200k tweets containing the hashtags `#surf`, `#windsurf` or `#kitesurf`.<br>This dataset has been built harvesting twitter every day during about 300 days. It is fully available in this [Github repository](https://github.com/holtzy/data_to_viz/tree/master/Example_dataset). Basically it looks like the table beside.
</div>
<div class = "col-md-4">
```{r, warning=FALSE, message=FALSE}
# Libraries
library(tidyverse)
library(viridis)
library(hrbrthemes)
library(kableExtra)
options(knitr.table.format = "html")

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/17_ListGPSCoordinates.csv", sep=",", header=T)

# show data
data %>% head(3) %>% kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
</div>
</div>






#Showing a few dots
***



#A dot density map
***


#Hexbin
***

```{r, warning=FALSE, message=FALSE, fig.alig="center"}
# Get the world polygon
library(mapdata)
world <- map_data("world")
 
data %>%
  filter(homecontinent=='Europe') %>%
  ggplot( aes(x=homelon, y=homelat)) + 
    geom_hex(bins=65) +
    theme_void() +
    xlim(-30, 70) +
    ylim(24, 72) +
    scale_fill_viridis(
      trans = "log", 
      name="Number of Tweet recorded in 8 months", 
      guide = guide_legend( keyheight = unit(3, units = "mm"), keywidth=unit(12, units = "mm"), label.position = "bottom", title.position = 'top', nrow=1) 
    )  +
    ggtitle( "Where people tweet about #Surf" ) +
    theme(
      legend.position = c(0.5, 0.09),
      text = element_text(color = "#22211d"),
      #plot.background = element_rect(fill = "#f5f5f2", color = NA), 
      #panel.background = element_rect(fill = "#f5f5f2", color = NA), 
      #legend.background = element_rect(fill = "#f5f5f2", color = NA),
      plot.title = element_text(size= 22, hjust=0.1, color = "#4e4d47", margin = margin(b = -0.1, t = 0.4, l = 2, unit = "cm")),
    )
```







```{r, warning=FALSE}

# Make the hexbin map with the geom_hex function
ggplot(data, aes(x=homelon, y=homelat)) + 
  geom_polygon(data = world, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
  geom_bin2d(bins=80) +
  theme_void() +
  scale_fill_viridis(
    trans = "log", 
    name="Number of Tweets", 
    guide = guide_legend( keyheight = unit(3, units = "mm"), keywidth=unit(12, units = "mm"), label.position = "bottom", title.position = 'top', nrow=1)
  ) +
  theme(
    legend.position = c(0.5, 0.09),
  ) +
  ylim(-70, 80)
```



#Choropleth
***


