---
myimage1: "../img/section/DensitySmall.png"
myimage2: "../img/section/HistogramSmall.png"
myimage3: "../img/section/Box1Small.png"
myimage4: "../img/section/ViolinSmall.png"
myimage5: "../img/section/JoyplotSmall.png"
mytitle: "Visualizing One Numeric and One categoric variable"
output:
  html_document:
      template: template_story.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: TRUE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---

<br><br><br>
<div class="mycontent">





<div class = "row">
<div class = "col-md-8">
<br><br>
This document gives a few suggestions to analyse a dataset composed by a numeric and a categoric variable.
<br><br>
On the [/r/samplesize](https://www.reddit.com/r/SampleSize/) thread of reddit, questions like *What probability would you assign to the phrase "Highly likely"* was asked. Results allow to understand how people perceive probability vocabulary.
<br><br>
**Disclaimer**: This idea originally comes from a publication of the [CIA](https://www.cia.gov/library/center-for-the-study-of-intelligence/csi-publications/books-and-monographs/psychology-of-intelligence-analysis/art15.html) which resulted in this [figure](https://www.cia.gov/library/center-for-the-study-of-intelligence/csi-publications/books-and-monographs/psychology-of-intelligence-analysis/fig18.gif/image.gif). Then, [Zoni Nation](https://github.com/zonination) cleaned the reddit dataset and built [graphics with R](https://github.com/zonination/perceptions). I heavily rely on his work in the folowing.

</div>
<div class = "col-md-4">
```{r, warning=FALSE, message=FALSE}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(kableExtra)
options(knitr.table.format = "html")
library(viridis)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/zonination/perceptions/master/probly.csv", header=TRUE, sep=",")
data <- data %>% 
  gather(key="text", value="value") %>%
  mutate(text = gsub("\\.", " ",text)) %>%
  mutate(value = round(as.numeric(value),0))
  

# show data
data %>% sample_n(8) %>% kable(row.names = FALSE) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
</div>
</div>










#Boxplot
***

The mos
```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE}
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=text, y=value, fill=text)) +
    geom_boxplot() +
    geom_jitter(color="grey", alpha=0.3, size=0.9) +
    scale_fill_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none"
    ) +
    coord_flip() +
    xlab("") +
    ylab("Assigned Probability (%)")
```


#Boxplot
***

The mos
```{r, fig.align='center', fig.height=8, fig.width=6, warning=FALSE}
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=text, y=value, fill=text, color=text)) +
    geom_violin(width=2.1, size=0.2) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none"
    ) +
    coord_flip() +
    xlab("") +
    ylab("Assigned Probability (%)")
```



#Density
***

If you have just a few group, you can compare them on the same plot.
```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE}
data %>%
  filter(text %in% c("Almost No Chance", "About Even", "Probable", "Almost Certainly")) %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=value, color=text, fill=text)) +
    geom_density(alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)")
```


However if you have more than ~4 groups this technique does not work: the graphic would become too cluttered. Thus it is a better practice to use small multiple.
```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE}
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=value, color=text, fill=text)) +
    geom_density(alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)") +
    facet_wrap(~text, scale="free_y")
```





#Histogram
***

```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE}
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=value, color=text, fill=text)) +
    geom_histogram(alpha=0.6, binwidth = 5) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)") +
    facet_wrap(~text, scale="free_y")
```


#Ridgeline plot
***
```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE}
library(ggridges)
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(y=text, x=value,  fill=text)) +
    geom_density_ridges(alpha=0.6, bandwidth=4) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)")
```
