---
myimage1: "Bar150.png"
mytitle: "Researchers network and migration flows"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_story.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: FALSE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---

<br><br>
<div class="mycontent">




Adjacency and incidence matrix provide relationship between several nodes. The information they contain can have different nature, thus this document will consider several examples:


- Relationships can be `directed` and `weighted`. Like the number of people migrating from one country to another. Data used comes from this [scientific publication](https://onlinelibrary.wiley.com/doi/abs/10.1111/imre.12327) from [Gui J. Abel](http://guyabel.com).
```{r, warning=FALSE, message=FALSE}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(circlize)
library(kableExtra)
options(knitr.table.format = "html")
library(viridis)
library(igraph)
library(ggraph)

# Load dataset from github
#data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/13_AdjacencyDirectedWeighted.csv", header=TRUE)
data <- read.table("../Example_dataset/13_AdjacencyDirectedWeighted.csv", header=TRUE)

# show data
data %>% head(3) %>% select(1:3) %>% kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

- Relationships can be `undirected` and `unweighted`. I will consider all the co-authors of a researcher and study who is connected through a common publication. Data have been retrieved using the [scholar](https://github.com/jkeirstead/scholar) package, the pipeline is describe in this [github repository](https://github.com/holtzy/Google-Scholar-Network). The result is an adjacency matrix with about 100 researchers, filled with 1 if they have published a paper together, 0 otherwise.

```{r, warning=FALSE, message=FALSE}
# Load data
#dataUU <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/13_AdjacencyUndirectedUnweighted.csv", header=TRUE)
dataUU <- read.table("../Example_dataset/13_AdjacencyUndirectedUnweighted.csv", header=TRUE)

# show data
dataUU %>% head(3) %>% select(1:4) %>% kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

- Relationships can also be `undirected` and `weighted`

- Relationships can also be `directed` and `unweighted`







# Chord diagram
***

Chord diagram is a good way to represent the migration flows. It works well if your data are directed and weighted like for migration flows between country.  

Disclaimer: this plot is made using the circlize library, and very strongly inspired from the [Migest package](https://github.com/cran/migest) from [Gui J. Abel](http://guyabel.com), who is also the author of the migration [dataset](https://www.oeaw.ac.at/fileadmin/subsites/Institute/VID/PDF/Publications/Working_Papers/WP2016_02.pdf) used here.

Since this kind of graphic is used to display flows, it can be applied only on network where connection are `weighted`. It does not work for the other example on authors connections.

```{r, fig.align="center", fig.width=7, fig.height=7}
# short names
colnames(data) <- c("Africa", "East Asia", "Europe", "Latin Ame.",   "North Ame.",   "Oceania", "South Asia", "South East Asia", "Soviet Union", "West.Asia")
rownames(data) <- colnames(data)

# I need a long format
data_long <- data %>%
  rownames_to_column %>%
  gather(key = 'key', value = 'value', -rowname)

# parameters
circos.clear()
circos.par(start.degree = 90, gap.degree = 4, track.margin = c(-0.1, 0.1), points.overflow.warning = FALSE)
par(mar = rep(0, 4))

# color palette
mycolor <- viridis(10, alpha = 1, begin = 0, end = 1, option = "D")
mycolor <- mycolor[sample(1:10)]

# Base plot
chordDiagram(
  x = data_long, 
  grid.col = mycolor,
  transparency = 0.25,
  directional = 1,
  direction.type = c("arrows", "diffHeight"), 
  diffHeight  = -0.04,
  annotationTrack = "grid", 
  annotationTrackHeight = c(0.05, 0.1),
  link.arr.type = "big.arrow", 
  link.sort = TRUE, 
  link.largest.ontop = TRUE)

# Add text and axis
circos.trackPlotRegion(
  track.index = 1, 
  bg.border = NA, 
  panel.fun = function(x, y) {
    
    xlim = get.cell.meta.data("xlim")
    sector.index = get.cell.meta.data("sector.index")
    
    # Add names to the sector. 
    circos.text(
      x = mean(xlim), 
      y = 3.2, 
      labels = sector.index, 
      facing = "bending", 
      cex = 0.8
      )

    # Add graduation on axis
    circos.axis(
      h = "top", 
      major.at = seq(from = 0, to = xlim[2], by = ifelse(test = xlim[2]>10, yes = 2, no = 1)), 
      minor.ticks = 1, 
      major.tick.percentage = 0.5,
      labels.niceFacing = FALSE)
  }
)
  
```
In my opinion this is a powerful way to display information. Major flows are easy to detect, like the migration from South Asia towars Westa Asia, or Africa to Europe. Moreover, for each continent it is quite easy to quantify the proportion of people leaving and arriving.

However chord diagram is not an usual way of displaying information. Thus, it is advised to give a good amount of explanation to educate your audience. A good way to do so is to draw just a few connections in a first step, before displaying the whole graphic. See this blog post by [Nadieh Bremer]() for more ideas on this topic.


# Sankey diagram
***

[Sankey diagram]() is another option to display weighted connection. Intead of displaying regions on a circle, they are duplicated and represented on both side of the graphic. Origin is usually on the left, destination on the right.
```{r, fig.align="center", fig.width=10, fig.height=7, warning=FALSE}
# Package
library(networkD3)

# I need a long format
data_long <- data %>%
  rownames_to_column %>%
  gather(key = 'key', value = 'value', -rowname) %>%
  filter(value > 0)
colnames(data_long) <- c("source", "target", "value")
data_long$target <- paste(data_long$target, " ", sep="")

# From these flows we need to create a node data frame: it lists every entities involved in the flow
nodes <- data.frame(name=c(as.character(data_long$source), as.character(data_long$target)) %>% unique())
 
# With networkD3, connection must be provided using id, not using real name like in the links dataframe.. So we need to reformat it.
data_long$IDsource=match(data_long$source, nodes$name)-1 
data_long$IDtarget=match(data_long$target, nodes$name)-1

# prepare colour scale
ColourScal ='d3.scaleOrdinal() .range(["#FDE725FF","#B4DE2CFF","#6DCD59FF","#35B779FF","#1F9E89FF","#26828EFF","#31688EFF","#3E4A89FF","#482878FF","#440154FF"])'

# Make the Network
sankeyNetwork(Links = data_long, Nodes = nodes,
                     Source = "IDsource", Target = "IDtarget",
                     Value = "value", NodeID = "name", 
                     sinksRight=FALSE, colourScale=ColourScal, nodeWidth=40, fontSize=13, nodePadding=20)
```





#Heatmap
***

```{r}

# with ggplot2 --> not very insightful if not clusterised
dataUU %>% 
  gather(key="to", value="value", -1) %>%
  ggplot( aes(x=to, y=from, fill=value)) +
    geom_tile()

tmp <- dataUU
rownames(tmp) <- tmp[,1]
tmp <- tmp[,-1]
tmp <- as.matrix(tmp)
tmp[which(is.na(tmp))] <- 0
heatmap(tmp, na.rm=TRUE)
summary(tmp)
```







# Network
***
Since an adjacency matrix is a network structure, it is possible to build a network graph. In my opinion, this makes more sense when the connection are `unweighted`, since drawing edges with different sizes tends to clutter the figure and make it unreadable.  

Thus, here is an application of this chart type to the coauthor network. When to people have published at least one scientific paper together, they are connected.
```{r}
# Transform the adjacency matrix in a long format
connect <- dataUU %>% 
  gather(key="to", value="value", -1) %>%
  na.omit()

# Number of connection per person
c( as.character(connect$from), as.character(connect$to)) %>%
  as.tibble() %>%
  group_by(value) %>%
  summarize(n=n()) -> coauth
colnames(coauth) <- c("name", "n")

# Create a graph object with igraph
mygraph <- graph_from_data_frame( connect, vertices = coauth )
plot(mygraph)

# Make the graph
ggraph(mygraph, layout="nicely") + 
  #geom_edge_density(edge_fill="#69b3a2") +
  geom_edge_link(edge_colour="black", edge_alpha=0.2, edge_width=0.3) +
  geom_node_point(aes(size=n, alpha=n)) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0), "null"),
    panel.spacing=unit(c(0,0,0,0), "null")
  ) 


```

Network chart is a whole field of study on its own and overtake the scope of this website from far. The biggest challenge is to position each node efficiently, to avoid crosses between edges. Several algorithms exist to tackle this issue and it is always a good practice to try several of them.

```{r, fig.align="center"}
# Make the graph
ggraph(mygraph, layout="auto") + 
  #geom_edge_density(edge_fill="#69b3a2") +
  geom_edge_link(edge_colour="black", edge_alpha=0.2, edge_width=0.3) +
  geom_node_point(aes(size=n, alpha=n)) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0), "null"),
    panel.spacing=unit(c(0,0,0,0), "null")
  ) 

ggraph(mygraph, layout="igraph", algorithm="kk") + 
  #geom_edge_density(edge_fill="#69b3a2") +
  geom_edge_link(edge_colour="black", edge_alpha=0.2, edge_width=0.3) +
  geom_node_point(aes(size=n, alpha=n)) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0), "null"),
    panel.spacing=unit(c(0,0,0,0), "null")
  ) 
```


# Chord diagram
***
Instead of using a custom algorithm to position each nodes, it is possible to place them al around a circule, a bit like for a chord diagram. But this kind of chart makes sense only if the order of nodes around the circule is carefully chosen, to avoid aving a fusball.
```{r}
#dataUU %>% hclust()
#help(hclust)
#dist(dataUU)
#cor(dataUU[,-1], use="complete.obs")
#summary(dataUU)
#as.dist(dataUU)
#dim(dataUU)
#hclust(d, method = "complete", members = NULL)

# Make the graph
ggraph(mygraph, layout="circle") + 
  #geom_edge_density(edge_fill="#69b3a2") +
  geom_edge_link(edge_colour="black", edge_alpha=0.2, edge_width=0.3) +
  geom_node_point(aes(size=n, alpha=n)) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0,0), "null"),
    panel.spacing=unit(c(0,0,0,0), "null")
  ) 
```


# Arc diagram
***
This is basically the same concept, but where nodes are displayed all along a line
```{r}
# Make the graph
ggraph(mygraph, layout="linear") + 
  geom_edge_arc(edge_colour="black", edge_alpha=0.2, edge_width=0.3, fold=TRUE) +
  geom_node_point(aes(size=n, alpha=n)) +
  geom_node_text(aes(label=name), angle=45, hjust=1) +
  theme_void() +
  theme(
    legend.position="none",
    plot.margin=unit(c(0,0,0.4,0), "null"),
    panel.spacing=unit(c(0,0,3.4,0), "null")
  ) 
```








 








</div>














