---
myimage1: "BubbleMap150.png"
pathSlug: "GPSCoordWithValue"
mytitle: "The biggest UK cities"
output:
  html_document:
      self_contained: false
      mathjax: default
      lib_dir: libs
      template: template_story.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: FALSE
      df_print: "paged"
      code_folding: "hide"
      includes:
        after_body: footer.html
---

<br><br><br><br>

<div class="mycontent">






<div class = "row">

<div class = "col-md-6">

<br>

This document gives a few suggestions to analyse a dataset composed by a set of geographic coordinates that have an associated numeric value.<br><br>It considers the population of 925 cities in the UK. This example dataset is provided in the [R maps library](https://github.com/adeckmyn/maps) and is available on this [Github repository](https://github.com/holtzy/data_to_viz/tree/master/Example_dataset). Basically it looks like the table beside.

</div>

<div class = "col-md-6">

```{r, warning=FALSE, message=FALSE}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(kableExtra)
options(knitr.table.format = "html")
library(viridis)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/18_ListGPSCoordinatesWithValue.csv", header=T)

# show data
data %>% head(5) %>% kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
</div>
</div>










# Bubble map
***

The go-to graphic in this kind of situation is probably the [bubble map](https://www.data-to-viz.com/graph/bubblemap.html). Basically, one circle is drawn per provided geographic position. The size of the bubble is proportional to its corresponding numeric value. Warning, remember that the bubble area must be proportional to the value, [not its radius](https://www.data-to-viz.com/caveat/radius_or_area.html).

```{r, message=FALSE, warning=FALSE, fig.align='center', fig.height=8, fig.width=7}
# Get the world polygon and extract UK
library(maps)
UK <- map_data("world") %>% filter(region=="UK")

# Easy to make it interactive!
library(plotly)

# plot
p=data %>%

  arrange(desc(pop)) %>%
  mutate( name=factor(name, unique(name))) %>%
  mutate( mytext=paste("City: ", name, "\n", "Population: ", pop, sep="")) %>%  # This prepare the text displayed on hover.

  # Makte the static plot calling this text:
  ggplot() +
    ggplot2::annotate("text", x = 1, y = 56.3, label="1000 biggest cities in the UK", colour = "black", size=4, alpha=1) +
    ggplot2::annotate("segment", x = -1, xend = 3, y = 56, yend = 56, colour = "black", size=0.2, alpha=1) +
    geom_polygon(data = UK, aes(x=long, y = lat, group = group), fill="grey", alpha=0.3) +
    geom_point(aes(x=long, y=lat, size=pop, color=pop, text=mytext, alpha=pop) ) +
    scale_size_continuous(range=c(0.3,12)) +
    scale_color_viridis(option="inferno", trans="log" ) +
    scale_alpha_continuous(trans="log") +
    theme_void() +
    ylim(50,59) +
    coord_map() +
    theme(
      legend.position = "none"
    )

p <- ggplotly(p, tooltip="text")
```


<center>
```{r, message=FALSE, warning=FALSE, fig.align='center', fig.height=9, fig.width=7, echo=FALSE}
p
```
</center>

*Note*: this map is `interactive`. Hover a bubble to get the city name and its population, Zoom on a specific part for more details.







