---
myimage1: "DensitySmall.png"
myimage2: "Box1Small.png"
myimage3: "ViolinSmall.png"
myimage4: "HistogramSmall.png"
pathSlug: "OneNumSevCatSubgroupSevObsPerGroup"
mytitle: "How much do people tip?"
output:
  html_document:
      self_contained: false
      mathjax: default
      lib_dir: libs
      template: template_story.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: FALSE
      df_print: "paged"
      code_folding: "hide"
      includes:
        after_body: footer.html
---

<br>
<div class="mycontent">


<div class = "row">
<div class = "col-md-8">
<br><br><br>
This document gives a few suggestions to analyse a dataset composed by a numeric variable measured on groups and subgroups, with several observations for each combinations.
<br><br>
The dataset used as an example describes the amount that restaurant staff receive in tips based on various indicators like the client sex, the day of the week and if the client smokes or not.
<br><br>
Data come from the [Seaborn](https://github.com/mwaskom/seaborn-data) Python library. A clean `.csv` version is available in the data-to-viz.com [github repository](https://github.com/holtzy/data_to_viz/tree/master/Example_dataset).
</div>

<div class = "col-md-4">
<br><br>
```{r, warning=FALSE, message=FALSE}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(kableExtra)
options(knitr.table.format = "html")
library(viridis)
library(ggrepel)
library(plotly)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/10_OneNumSevCatSubgroupsSevObs.csv", header=T, sep=",") %>%
  mutate(tip = round(tip/total_bill*100, 1))

# show data
data %>% head(6) %>% select(tip, sex, day, smoker) %>% kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```
</div>
</div>





#Grouped boxplot
***
The most common way to represent that kind of dataset is probably the grouped [boxplot](https://www.data-to-viz.com/caveat/boxplot.html). Each combination of group is represented by a box. This allows to quickly understand what is the distribution of the numeric variable for each combination.

<br>

Here, it looks like there is not much difference in tip values from one day to the other in average, except a slight increase on sunday. Moreover, it looks like females tend to tip more than males on friday. Note that individual data points are presented using `jittering`, what allows to detect more particular pattern and assess the sample size of each group.
```{r, fig.width=8, fig.height=6, fig.align="center", warning=FALSE}
# Counts the number of value per group and subgroup
counts = data %>%
  group_by(day, sex) %>%
  summarize(
    n=n(),
    median=median(tip)
  )

# Grouped
data %>%
  mutate(day = fct_reorder(day, tip)) %>%
  mutate(day = factor(day, levels=c("Thur", "Fri", "Sat", "Sun"))) %>%
  ggplot(aes(fill=sex, y=tip, x=day)) +
    geom_boxplot(position=position_dodge2(preserve = "total"), alpha=0.5, outlier.colour="transparent", varwidth = TRUE) +
    geom_point(color="grey", size=1, width=0.1, position=position_jitterdodge() , alpha=0.4) +
    scale_fill_viridis(discrete=T, name="") +
    geom_text(data=counts, aes(label=paste0("n: ",n), y=median-2), position=position_dodge(1), hjust=0.5) +
    theme_ipsum()  +
    xlab("") +
    ylab("Tip (%)") +
    ylim(0,35)
```

In the above chart categories are grouped by day. It is possible to build the same kind of visualization, grouping by Sex instead. In the first case, it is easy to compare the behavior or each sex day by day. On the second one, the idea is more to compare the difference by day for each sex.


```{r, fig.width=8, fig.height=6, fig.align="center", warning=FALSE}
# Grouped
data %>%
  mutate(day = fct_reorder(day, tip)) %>%
  mutate(day = factor(day, levels=c("Thur", "Fri", "Sat", "Sun"))) %>%
  ggplot(aes(fill=day, y=tip, x=sex)) +
    geom_boxplot(position="dodge", alpha=0.5, outlier.colour="transparent") +
    geom_point(color="grey", size=1, width=0.1, position=position_jitterdodge() , alpha=0.4) +
    scale_fill_viridis(discrete=T, name="") +
    theme_ipsum()  +
    xlab("") +
    ylab("Tip (%)") +
    ylim(0,40)
```




#Violin
***

The [violin plot](https://www.data-to-viz.com/graph/violin.html) can be used exactly in the same situation that the [boxplot](https://www.data-to-viz.com/caveat/boxplot.html). Using the same grouping technique, we got here a grouped violin plot.
```{r, fig.width=8, fig.height=6, fig.align="center", warning=FALSE}
# Grouped
data %>%
  mutate(day = fct_reorder(day, tip)) %>%
  mutate(day = factor(day, levels=c("Thur", "Fri", "Sat", "Sun"))) %>%
  ggplot(aes(fill=sex, y=tip, x=day)) +
    geom_violin(position="dodge", alpha=0.5, outlier.colour="transparent") +
    scale_fill_viridis(discrete=T, name="") +
    theme_ipsum()  +
    xlab("") +
    ylab("Tip (%)") +
    ylim(0,40)
```






#Histogram
***
Small multiple is a powerful technique that can be used with that kind of data. Here, each combination is represented in a part of the grid, allowing a quick comparison between them. [Histograms](https://www.data-to-viz.com/graph/histogram.html) are used here, but the same result could be obtained with [density](https://www.data-to-viz.com/graph/density.html) plot.
```{r, fig.width=8, fig.height=6, fig.align="center", warning=FALSE}
# Grouped
data %>%
  mutate(day = fct_reorder(day, tip)) %>%
  mutate(day = factor(day, levels=c("Thur", "Fri", "Sat", "Sun"))) %>%
  ggplot(aes(x=tip)) +
    geom_histogram(bins=20, fill="#69b3a2", color="white") +
    facet_grid(sex~time) +
    theme_ipsum()  +
    xlab("") +
    ylab("Tip")
```









