---
myimage1: "DensitySmall.png"
myimage2: ".Box1Small.png"
myimage3: "ViolinSmall.png"
myimage4: "JoyplotSmall.png"
myimage5: "HeatmapSmall.png"
myimage6: "DendrogramSmall.png"
mydisqus: "SeveralNum"
mytitle: "Eleven features for 32 cars"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_story.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: FALSE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---

<br><br>
<div class="mycontent">


This page provides a few hints to visualize a dataset composed of several numeric variables. As an example the famous mtcars dataset will be considered. It provides several features like the number of cylinders, the gross horsepower, the weight etc. for about  32 car models.


```{r, warning=FALSE, message=FALSE}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(DT)
library(plotly)
library(dendextend)
library(car)
library(FactoMineR)
library(kableExtra)
options(knitr.table.format = "html")

# This dataset is available in R by default, and on the datatoviz github repo
data <- mtcars

# Save it at .csv for the github repo
#write.csv(mtcars, file="../Example_dataset/6_SeveralNum.csv", quote=F)

# show data
data %>% head(6) %>% kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

#- Check distribution
***

In my opinion, the first thing to do when you have several numeric variables is to observe their distribution one by one. This can be done using a [violin plot](), a [boxplot]() or a [ridgeline plot]() if your variables are all on the same scale. In the case of the `mtcars` dataset the variables are completely different one to each other so it make more sense to make an histogram for each of them:

```{r, warning=FALSE, message=FALSE, fig.width=10, fig.height=6, fig.align="center"}

# Keep the numeric variables of the mtcars dataset
data <- mtcars %>% select( disp, drat, hp, mpg, qsec, wt)

# Show the histogram of these variables
data %>%
  as.tibble() %>%
  gather(variable, value) %>%
  ggplot( aes(x=value) ) +
    geom_histogram( fill= "#69b3a2") +
    facet_wrap(~variable, scale="free") +
    theme_ipsum()
```






#- Correlogram
***

```{r, fig.align="center", fig.width=7, fig.height=7}
scatterplotMatrix(~mpg+disp+drat+hp+qsec+wt, data=data , reg.line="" , col="#69b3a2", id.col="#69b3a2", smoother="" , smoother.args=list(col="grey") , cex=1.5 , pch=20 )
```

#- Dendrogram
***

```{r, fig.align="center", fig.width=7}
# Clusterisation using 3 variables
data %>% dist() %>% hclust() %>% as.dendrogram() -> dend

# Color in function of the cluster
par(mar=c(1,1,1,7))
dend %>%
  set("labels_col", value = c("#69b3a2", "#404080", "orange"), k=3) %>%
  set("branches_k_color", value = c("#69b3a2", "#404080", "orange"), k = 3) %>%
  plot(horiz=TRUE, axes=FALSE)
abline(v = 350, lty = 2)
 
```

#- Heatmap
***

```{r, fig.align="center", fig.width=8}
library(d3heatmap)
d3heatmap(data, k_row = 4, k_col = 2, scale = "column")
```


#- PCA
***

```{r, fig.align="center", fig.width=7}
myPCA <- PCA(data, scale.unit=TRUE, graph=F)

myPCA$ind %>%
  as.data.frame() %>%
  mutate(name=rownames(.)) %>%
  ggplot( aes(x=coord.Dim.1, y=coord.Dim.2, label=name)) +
    geom_point( color="#69b3a2") +
    theme_ipsum() +
    geom_label(color="#69b3a2")

#plot.PCA(myPCA, axes=c(1, 2), choix="var")

```



#- Getting a correlation matrix
***

It is of importance to note that this kind of dataset can be converted to a correlation matrix that is an adjacency matrix. Indeed, we can compute the correlation between each pair of variable or each pair of entities of the dataset and try to visualize this new dataset. But this is a new story: how to visualize an adjacency matrix.
