---
myimage1: "Line150.png"
myimage2: "Area150.png"

mytitle: "Evolution of the bitcoin price"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_story.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: FALSE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---

<br>
<div class="mycontent">




<div class = "row">
<div class = "col-md-8">
<br><br><br>
This document gives a few suggestions to analyse a dataset composed by two ordered numeric variables. It considers the evolution of the [bitcoin](https://en.wikipedia.org/wiki/Bitcoin) price between April 2013 and April 2018.
<br><br>
This dataset has been built using the [crypto](https://github.com/JesseVent/crypto) R package that allows to access the [ CoinMarketCap](https://coinmarketcap.com) website. The first column, `date`, represents an ordered numeric variable. The second, `value` gives the bitcoin price. This dataset is available on [github](https://github.com/holtzy/data_to_viz/tree/master/Example_dataset).
</div>
<div class = "col-md-4">
<br>
```{r, warning=FALSE, message=FALSE}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(DT)
library(plotly)
library(kableExtra)
options(knitr.table.format = "html")

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/2_TwoNumOrdered.csv", header=T)
data$date <- as.Date(data$date)

# Show long format
data %>%
  head(5) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```



</div>
</div>








#- Line plot
***
The most comon way to represent that kind of dataset is probably to produce a [line plot](). It allows to give a good overview of the bitcoin price on the period, reaching a value of 20,000 $ in december 2017.

```{r, fig.align="center", fig.width=10}
data %>%
  ggplot( aes(x=date, y=value)) +
    geom_line(color="#69b3a2") +
    ggtitle("Evolution of Bitcoin price") +
    ylab("bitcoin price ($)") +
    theme_ipsum()
```





#- Area chart
***
Even if the line chart is a very good way to convey the information, when only one variable is represented like here, the plot can appear to be slightly empty. Thus, an interesting alternative is the [area chart](). It is basically the same thing, except that the area between the X axis and the line is filled.

```{r, fig.align="center", fig.width=10}
data %>%
  ggplot( aes(x=date, y=value)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ggtitle("Evolution of Bitcoin price") +
    ylab("bitcoin price ($)") +
    theme_ipsum()
```

Note that using the same color for the filled area and the line is often a good looking choice, with a bit more transparency for the filled area.


#- Interactivity
***
Using interactivity gives a great added value to your line or area chart. Indeed, it is very useful to ba able to zoom on a specific time slot of interest on the graphic. For instance, have a look to what happened in 2014: bitcoin already experienced a huge evolution, comparable to 2017 in terme of relative evolution, but not mediatised at all.
```{r, fig.align="center", fig.width=10}
p <- data %>%
  ggplot( aes(x=date, y=value)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ggtitle("Evolution of Bitcoin price") +
    ylab("bitcoin price ($)") +
    theme_ipsum()
ggplotly(p)
```




#- Few points? Use connected scatter
***
If you have just a few point in your dataset, you probably want to use a [connected scatterplot]() instead. it is basically the same thing where each individual point is represented. It greatly helps to understand when observation have been made. Let's consider the last 10 observation of the bitcoin dataset:
```{r, fig.align="center", fig.width=10}
data %>%
  tail(10) %>%
  ggplot( aes(x=date, y=value)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    geom_point() +
    ggtitle("Evolution of Bitcoin price") +
    ylab("bitcoin price ($)") +
    theme_ipsum()
```




#- To cut or not to cut?
***
The previous chart can be a bit frustrating. It is indeed hard to study the evolution of the bitcoin in this graphic since the price ranges between 7,500 and 10,000 dollars in this period, when the Y axis ranges between 0 and 10,000. In this case, it is a good practice to cut the Y axis to zoom on the variation. This subject is subject to many debates in the dataviz community and you can read more about it in the [dedicated page]().
```{r, fig.align="center", fig.width=10}
data %>%
  tail(10) %>%
  ggplot( aes(x=date, y=value)) +
    geom_ribbon(aes(ymin=8000, ymax=value),  fill="#69b3a2", color="transparent", alpha=0.5) +
    geom_line(color="#69b3a2") +
    geom_point() +
    ggtitle("Evolution of Bitcoin price") +
    ylab("bitcoin price ($)") +
    theme_ipsum()
```


#- Comparing to a limit
***
Sometimes one can be interested in comparing the value to a specific threshold. In this case, you can fill the area depending on this threshold, with 2 different colors depending if the value is over or below the threshold:
```{r, fig.align="center", fig.width=10}
data %>%
  tail(300) %>%
  mutate( mycolor=ifelse(value>7500, "yes", "no")) %>%
  ggplot( aes(x=date, y=value)) +
    geom_ribbon(aes(ymin=7500, ymax=value, fill=mycolor), color="black", alpha=0.5) +
    scale_fill_manual(values=c("#69b3a2","#271569")) +
    ggtitle("Evolution of Bitcoin price") +
    ylab("bitcoin price ($)") +
    theme_ipsum() +
    theme(legend.position="none")
```

Note: this graphic is imperfect and must be improved (don't understand the behavior of `geom_ribbon`)
