---
myimage1: "../img/mistake/circular_bar_yaxis.png"
mytitle: "Circular barplot and distortion"
mydisqus: "circular_bar_yaxis"
output:
  html_document:
      template: template_caveat.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: TRUE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---

<br><br>
<div class="mycontent">



#The issue
***
[Circular barplots](https://www.data-to-viz.com/graph/circularbarplot.html) are a good alternative to classic [barplots](https://www.data-to-viz.com/graph/circularbarplot.html). It is basically the same thing with each bar displayed along a circle instead of a line, resulting in a more eye-catching figure.

<br>

However, circular barplots are `less efficient` to represent data `accurately`. Indeed, it is harder to compare bar lenghts since they are not on the same line anymore. Moreover, the basis of the bar is often thiner than the top part. This results in a bias: group with large values appear even larger than they are in reality.

<br>

This bias is even more important if:

- *left* - the number of bar is low.
- *right* - the inner part of the circle is too small.

```{r, warning=FALSE, message=FALSE, fig.align="center", fig.width=10, fig.height=5}
# package
library(tidyverse)
library(viridis)
library(patchwork)

# Create dataset
data=data.frame(
  individual=paste( "Mister ", seq(1,9), sep=""),
  group=c( rep('A', 3), rep('C', 3), rep('D', 3)) ,
  value=sample( seq(10,100), 9, replace=T)
)
data = data %>% arrange(group, value)

# Set a number of 'empty bar' to add at the end of each group
empty_bar=1
to_add = data.frame( matrix(NA, empty_bar*nlevels(data$group), ncol(data)) )
colnames(to_add) = colnames(data)
to_add$group=rep(levels(data$group), each=empty_bar)
data=rbind(data, to_add)
data=data %>% arrange(group)
data$id=seq(1, nrow(data))
 
# Get the name and the y position of each label
label_data=data
number_of_bar=nrow(label_data)
angle= 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust<-ifelse( angle < -90, 1, 0)
label_data$angle<-ifelse(angle < -90, angle+180, angle)
 

# Make the plot
p1 = ggplot(data, aes(x=as.factor(id), y=value, fill=group)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  
  geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=0.5) +
  scale_fill_viridis(discrete=TRUE) +
  ylim(-80,120) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar() + 
  geom_text(data=label_data, aes(x=id, y=value+10, label=individual, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=2.5, angle= label_data$angle, inherit.aes = FALSE ) 
  


# Create dataset 2
data=data.frame(
  individual=paste( "Mister ", seq(1,60), sep=""),
  group=c( rep('A', 20), rep('C', 24), rep('D', 16)) ,
  value=sample( seq(10,100), 60, replace=T)
)
data = data %>% arrange(group, value)

# Set a number of 'empty bar' to add at the end of each group
empty_bar=1
to_add = data.frame( matrix(NA, empty_bar*nlevels(data$group), ncol(data)) )
colnames(to_add) = colnames(data)
to_add$group=rep(levels(data$group), each=empty_bar)
data=rbind(data, to_add)
data=data %>% arrange(group)
data$id=seq(1, nrow(data))
 
# Get the name and the y position of each label
label_data=data
number_of_bar=nrow(label_data)
angle= 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust<-ifelse( angle < -90, 1, 0)
label_data$angle<-ifelse(angle < -90, angle+180, angle)
 

# Make the plot
p2 <- ggplot(data, aes(x=as.factor(id), y=value, fill=group)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  
  geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=0.5) +
  scale_fill_viridis(discrete=TRUE) +
  ylim(-10,120) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar() + 
  geom_text(data=label_data, aes(x=id, y=value+10, label=individual, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=2.5, angle= label_data$angle, inherit.aes = FALSE ) 
  

p1 + p2
```






#Solving the issue
***
Circular barcharts can be a good option in very specific conditions. You need: 

- a high number of bar
- a clear pattern in the data: if all group have almost the same value, circular version is not accurate enough for comparison.
- to keep a large inner circle

```{r, warning=FALSE, message=FALSE, fig.align="center"}
# package
library(tidyverse)
library(viridis)

# Create dataset
data=data.frame(
  individual=paste( "Mister ", seq(1,60), sep=""),
  group=c( rep('A', 20), rep('C', 24), rep('D', 16)) ,
  value=sample( seq(10,100), 60, replace=T)
)
data = data %>% arrange(group, value)

# Set a number of 'empty bar' to add at the end of each group
empty_bar=1
to_add = data.frame( matrix(NA, empty_bar*nlevels(data$group), ncol(data)) )
colnames(to_add) = colnames(data)
to_add$group=rep(levels(data$group), each=empty_bar)
data=rbind(data, to_add)
data=data %>% arrange(group)
data$id=seq(1, nrow(data))
 
# Get the name and the y position of each label
label_data=data
number_of_bar=nrow(label_data)
angle= 90 - 360 * (label_data$id-0.5) /number_of_bar     # I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
label_data$hjust<-ifelse( angle < -90, 1, 0)
label_data$angle<-ifelse(angle < -90, angle+180, angle)
 

# Make the plot
p = ggplot(data, aes(x=as.factor(id), y=value, fill=group)) +       # Note that id is a factor. If x is numeric, there is some space between the first bar
  
  geom_bar(aes(x=as.factor(id), y=value, fill=group), stat="identity", alpha=0.5) +
  scale_fill_viridis(discrete=TRUE) +
  ylim(-80,120) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1,4), "cm") 
  ) +
  coord_polar() + 
  geom_text(data=label_data, aes(x=id, y=value+10, label=individual, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=2.5, angle= label_data$angle, inherit.aes = FALSE ) 
  
p
```


*Note*: don't forget to add a `legend` or `annotation` for the colors and a `scale` for the bar height. See more in the [circular barplot section](file:///Users/y.holtz/Documents/data_to_viz/graph/circularbarplot.html).









#Comments
***
Any thoughts on this? Found any mistake? Disagree? Please drop me a word on [twitter](https://twitter.com/R_Graph_Gallery) or in the comment section below:
<br>




</div>














