---
myimage1: "../img/mistake/ungrouped_grouped_barplot.png"
mytitle: "Grouped barplot must be grouped"
mydisqus: "grouped_bar"
output:
  html_document:
      template: template_caveat.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: TRUE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---

<br><br>
<div class="mycontent">



#The issue
***
A [barplot](https://www.data-to-viz.com/graph/barplot.html) displays the value of a numeric variable for several entities. These entities can be grouped using a categoric variable, resulting in a `grouped barplot`.

<br>

When building a grouped barplot, make sure that your bar are indeed grouped: bars must be closer into each group than between group. In the example below, bars are not grouped, making hard to distinguish groups.
```{r, warning=FALSE, message=FALSE, fig.align="center"}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(babynames)
library(viridis)

# Load dataset from github
data <- babynames %>% 
  filter(name %in% c("Anna", "Mary")) %>%
  filter(sex=="F")

# A grouped barplot
data  %>% 
  filter(year %in% c(1950, 1960, 1970, 1980, 1990, 2000)) %>%
  mutate(year=as.factor(year)) %>%
  mutate( nameYear = paste(year, name, sep=" - ")) %>%
  ggplot( aes(x=as.factor(nameYear), y=n, fill=name)) +
    geom_bar(stat="identity") +
    scale_fill_viridis(discrete=TRUE, name="") +
    theme_ipsum() +
    ylab("Number of baby") +
    xlab("") +
    theme(
      axis.text.x = element_text(angle=60, hjust=1)
    )
```



#Solving the issue
***
The workaround is pretty simple, just group your bars:
```{r, warning=FALSE, message=FALSE, fig.align="center"}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(babynames)
library(viridis)

# Load dataset from github
data <- babynames %>% 
  filter(name %in% c("Anna", "Mary")) %>%
  filter(sex=="F")

# A grouped barplot
data  %>% 
  filter(year %in% c(1950, 1960, 1970, 1980, 1990, 2000)) %>%
  mutate(year=as.factor(year)) %>%
  ggplot( aes(x=year, y=n, fill=name)) +
    geom_bar(stat="identity", position="dodge") +
    scale_fill_viridis(discrete=TRUE, name="") +
    theme_ipsum() +
    ylab("Number of baby")
```








#Comments
***
Any thoughts on this? Found any mistake? Disagree? Please drop me a word on [twitter](https://twitter.com/R_Graph_Gallery) or in the comment section below:
<br>




</div>














