---
myimage1: "../img/mistake/hard_label.png"
mytitle: "A note on long labels"
pathSlug: "hard_label"
output:
  html_document:
      template: template_caveat.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_section: FALSE
      df_print: "paged"
      code_folding: "hide"
      includes:
        after_body: footer.html
---

<br><br>
<div class="mycontent">



# The issue
***
Most of the time, [barplot](https://www.data_to_viz.com/graph/barplot.html) or [lollipop](https://www.data_to_viz.com/graph/lollipop.html) plots are plotted vertically with the Y-axis representing the value of the numeric variable. If your labels on the X-axis are long, they need to be `rotated` in order not to overlap.
<br>

As a result, these labels become hard to read:
```{r, warning=FALSE, message=FALSE, fig.align="center"}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(kableExtra)
options(knitr.table.format = "html")

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/7_OneCatOneNum.csv", header=TRUE, sep=",")

# Barplot
data %>%
  filter(!is.na(Value)) %>%
  arrange(Value) %>%
  tail(20) %>%
  mutate(Country=factor(Country, Country)) %>%
  ggplot( aes(x=Country, y=Value) ) +
    geom_bar(stat="identity", fill="#69b3a2") +
    theme_ipsum() +
    theme(
      panel.grid.minor.x = element_blank(),
      panel.grid.major.x = element_blank(),
      legend.position="none",
      axis.text.x = element_text(angle = 80, hjust=1)
    ) +
    xlab("") +
    ylab("Weapon quantity (SIPRI trend-indicator value)")
```
*Note*: this barplot shows the quantity of weapons exported by the top 20 largest exporters in 2017, [read more](https://www.data-to-viz.com/story/OneNumOneCat.html).



# Solving the issue
***
The workaround is pretty simple, why not considering an `horizontal version` of the chart?
```{r, warning=FALSE, message=FALSE, fig.align="center"}
# Barplot
data %>%
  filter(!is.na(Value)) %>%
  arrange(Value) %>%
  tail(20) %>%
  mutate(Country=factor(Country, Country)) %>%
  ggplot( aes(x=Country, y=Value) ) +
    geom_bar(stat="identity", fill="#69b3a2") +
    theme_ipsum() +
    theme(
      panel.grid.minor.y = element_blank(),
      panel.grid.major.y = element_blank(),
      legend.position="none"
    ) +
    xlab("") +
    ylab("Weapon quantity (SIPRI trend-indicator value)") +
    coord_flip()
```





# Warning
***
Note that the horizontal version is not always an option though. If you categorical variable has a natural order, it is better to stick to the vertical version. It happens when:

- you're represented `time series`: time must be represented on the X axis by convention. Not doing so could mislead your audience.
- you have an `ordinal variable` like age range.



# Going further
***

- Bar Charts: vertical or horizontal? by [Ann K. Emery](http://annkemery.com/vertical-or-horizontal/)


<br>

</div>