---
myimage1: "../img/mistake/multi_distribution.png"
mytitle: "Too many distributions"
mydisqus: "multi_distribution"
output:
  html_document:
      template: template_caveat.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: TRUE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---

<br><br>
<div class="mycontent">


Comparing the distribution of several numeric variables is a common task in dataviz. Distribution of a variable can be represented using a [histogram](http://www.data-to-viz.com/graph/histogram.html) or a [density chart](http://www.data-to-viz.com/graph/density.html), and it is very tempting to represent many distribution on the same axis.  

Here is an example showing how people perceive probability vocabulary. On the [/r/samplesize](https://www.reddit.com/r/SampleSize/) thread of reddit, questions like *What probability would you assign to the phrase "Highly likely"* were asked. Here is the distribution of the score given by people to each question:


```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE, message=FALSE}
#library
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(patchwork)


# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/zonination/perceptions/master/probly.csv", header=TRUE, sep=",")
data <- data %>% 
  gather(key="text", value="value") %>%
  mutate(text = gsub("\\.", " ",text)) %>%
  mutate(value = round(as.numeric(value),0))

# A dataframe for annotations
annot <- data.frame(
  text = c("Almost No Chance", "About Even", "Probable", "Almost Certainly"),
  x = c(5, 53, 65, 79),
  y = c(0.15, 0.4, 0.06, 0.1)
)

# Plot
data %>%
  filter(text %in% c("Almost No Chance", "About Even", "Probable", "Almost Certainly")) %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=value, color=text, fill=text)) +
    geom_density(alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    geom_text( data=annot, aes(x=x, y=y, label=text, color=text), hjust=0, size=4.5) +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)")
```

In this case, the figure is quite neat. People gives a probability between 0 and 20 for the sentence "Almost No chance", and between 75 and 100% for "Almost Certainly". Let's check what happens with more groups:

```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE}
# Plot
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=value, color=text, fill=text)) +
    geom_density(alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)")
```

Now the figure is way to cluttered and it is impossible to distinguish groups: there are too many distributions represented on the same graphic. How to avoid that?


# Workaround {.tabset}
***
Several workarounds exist and are presented in detail in this [dedicated page](http://www.data-to-viz.com/story/OneNumOneCatSeveralObs.html). Here is just an overview of the possibilities:

```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE}
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=text, y=value, fill=text)) +
    geom_boxplot() +
    geom_jitter(color="grey", alpha=0.3, size=0.9) +
    scale_fill_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none"
    ) +
    coord_flip() +
    xlab("") +
    ylab("Assigned Probability (%)")
```

## Violin
```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE}
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=text, y=value, fill=text, color=text)) +
    geom_violin(width=2.1, size=0.2) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none"
    ) +
    coord_flip() +
    xlab("") +
    ylab("Assigned Probability (%)")
```

## Ridgeline
```{r, fig.align='center', fig.height=6, fig.width=8, warning=FALSE}
library(ggridges)
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(y=text, x=value,  fill=text)) +
    geom_density_ridges(alpha=0.6, bandwidth=4) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)")
```


## Faceting
```{r, fig.align='center', fig.height=7, fig.width=8, warning=FALSE}
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(x=value, color=text, fill=text)) +
    geom_density(alpha=0.6) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)") +
    facet_wrap(~text, scale="free_y")
```







#Going further
***

**Disclaimer**: This idea originally comes from a publication of the [CIA](https://www.cia.gov/library/center-for-the-study-of-intelligence/csi-publications/books-and-monographs/psychology-of-intelligence-analysis/art15.html) which resulted in this [figure](https://www.cia.gov/library/center-for-the-study-of-intelligence/csi-publications/books-and-monographs/psychology-of-intelligence-analysis/fig18.gif/image.gif). Then, [Zoni Nation](https://github.com/zonination) cleaned the reddit dataset and built [graphics with R](https://github.com/zonination/perceptions). I heavily rely on his work in this post.


**Other link**:

- Doing boxplots in [R](https://www.r-graph-gallery.com/boxplot/) and [Python](https://python-graph-gallery.com/boxplot/)
- Doing violin plots in [R](https://www.r-graph-gallery.com/violin/) and [Python](https://python-graph-gallery.com/violin/)


#Comments
***
Any thoughts on this? Found any mistake? Disagree? Please drop me a word on [twitter](https://twitter.com/R_Graph_Gallery) or in the comment section below:
<br>


</div>














