---
myimage1: "../img/mistake/spaghetti.png"
mytitle: "The Spaghetti plot"
mydisqus: "spaghetti"
output:
  html_document:
      template: template_caveat.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_section: FALSE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---

<br><br>
<div class="mycontent">

A Spaghetti plot is a [line plot](https://www.data-to-viz.com/graph/line.html) with many lines displayed together. With more than a few (~5?) groups this kind of graphic gets really hard to read, and thus provides little insight about the data. Let's make an example with the [evolution of baby names](http://www.data-to-viz.com/story/OneCatSevOrderedNum.html) in the US from 1880 to 2015.

```{r, warning=FALSE, message=FALSE, fig.align="center", fig.width=10}
# Libraries
library(tidyverse)
library(hrbrthemes)
library(kableExtra)
options(knitr.table.format = "html")
library(babynames)
library(streamgraph)
library(viridis)
library(DT)
library(plotly)

# Load dataset from github
data <- babynames %>% 
  filter(name %in% c("Mary","Emma", "Ida", "Ashley", "Amanda", "Jessica",    "Patricia", "Linda", "Deborah",   "Dorothy", "Betty", "Helen")) %>%
  filter(sex=="F")

# Plot
data %>%
  ggplot( aes(x=year, y=n, group=name, color=name)) +
    geom_line() +
    scale_color_viridis(discrete = TRUE) +
    theme(
      legend.position="none",
      plot.title = element_text(size=14)
    ) +
    ggtitle("A spaghetti chart of baby names popularity") +
    theme_ipsum()
```

It is very hard to follow a line to understand the evolution of a specific name's popularity. Plus, even if you manage to follow a line, you then need to link it with the legend which is even harder. Let's try to find a few workarounds to improve this graphic.



#Target a specific group
***
Let’s say you plot many groups, but the actual reason for that is to explain the feature of one particular group compared to the others. Then a good workaround is to highlight this group: make it appear different, and give it a proper annotation. Here, the evolution of Amanda's popularity is obvious. Leaving the other lines is important since it allows you to compare Amanda to all other names.

```{r, fig.align="center", fig.width=10}
data %>%
  mutate( highlight=ifelse(name=="Amanda", "Amanda", "Other")) %>%
  ggplot( aes(x=year, y=n, group=name, color=highlight, size=highlight)) +
    geom_line() +
    scale_color_manual(values = c("#69b3a2", "lightgrey")) +
    scale_size_manual(values=c(1.5,0.2)) +
    theme(legend.position="none") +
    ggtitle("Popularity of American names in the previous 30 years") +
    theme_ipsum() +
    geom_label( x=1990, y=55000, label="Amanda reached 3550\nbabies in 1970", size=4, color="#69b3a2") +
    theme(
      legend.position="none",
      plot.title = element_text(size=14)
)
```



#Use small multiples
***
[Area charts](https://www.data-to-viz.com/graph/area.html) can be used to give a more general overview of the dataset, especially when used in combination with small multiples. In the following chart, it is easy to get a glimpse of the evolution of any name:
```{r, fig.align="center", fig.width=8, fig.height=8}
data %>%
  ggplot( aes(x=year, y=n, group=name, fill=name)) +
    geom_area() +
    scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Popularity of American names in the previous 30 years") +
    theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      plot.title = element_text(size=14)
    ) +
    facet_wrap(~name)
```
For instance, `Linda` was a really popular name for a really short period of time. On another hand, Ida has never been very popular, but was used a little during several decades.



#Combine approaches
***
Note that if you want to compare the evolution of each line compared to the others, you can combine both approaches:
```{r, fig.align="center", fig.width=8, fig.height=8}
tmp <- data %>%
  mutate(name2=name)

tmp %>%
  ggplot( aes(x=year, y=n)) +
    geom_line( data=tmp %>% dplyr::select(-name), aes(group=name2), color="grey", size=0.5, alpha=0.5) +
    geom_line( aes(color=name), color="#69b3a2", size=1.2 )+
    scale_color_viridis(discrete = TRUE) +
    theme_ipsum() +
    theme(
      legend.position="none",
      plot.title = element_text(size=14),
      panel.grid = element_blank()
    ) +
    ggtitle("A spaghetti chart of baby names popularity") +
    facet_wrap(~name)

```


#Going further
***

- strategies for avoiding the spaghetti graph, by [storytelling with data](http://www.storytellingwithdata.com/blog/2013/03/avoiding-spaghetti-graph)
- doing line plots in [R](https://www.r-graph-gallery.com/line-plot/) and [Python](https://python-graph-gallery.com/line-chart/)


#Comments
***
Any thoughts on this? Found any mistake? Disagree? Please drop me a word on [twitter](https://twitter.com/R_Graph_Gallery) or in the comment section below:
<br>



</div>









